### 2020-09-16 paper/perfectGraphs.tex
Fixes

Given a graph $G$, let us consider a problem of coloring it using as few colors as possible. If $G$ contains a clique $K$ as a subgraph, we must use at least $|V(K)|$ colors to color it. This gives us a lower bound for a chromatic number $\chi(G)$ -- it is always greater or equal to the cardinality of the largest clique $\omega(G)$. The reverse is not always true, in fact we can construct a graph with no triangle and requiring arbitrarily large numbers of colors (e.g. construction by Mycielski \cite{Mycielski1955}).

Do graphs that admit coloring using only $\omega(G)$ color are "simpler" to further analyze? Not necessarily so. Given a graph $G = (V, E)$, $|V| = n$, let us construct a graph $G'$ as the union of $G$ and a clique $K_n$. We can see that indeed $\chi(G') = \omega(G') = n$, but it gives us no indication of the structure of $G$ or $G'$. This leads us to the hereditary definition of \emph{perfect graphs}.

@@

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003}, algorithms to solve semidefinite programs grew out of the theory of perfect graphs. We will take a look at semidefinite programs and at perfect graph's relation to Shannon capacity in \Cref{sec:ShannonCapacity}.

@@

\begin{theorem}[Perfect graph theorem from \cite{LL72}]
	\label{thm:pgt}
	A graph is perfect if and only if its complement graph is also perfect.
\end{theorem}

@@

Then, since $\alpha(H) = \omega(\overline{H})$ and $\omega(H) = \alpha(\overline{H})$ \Cref{thm:omegaalpha} implies \Cref{thm:pgt}.

@@

Berge graph recognition algorithm (later called CCLSV from the names of its authors) could be divided into two parts: first we check if either $G$ or $\overline{G}$ contain any of a number of simple forbidden structures (\Cref{SimpleStructures}). If they do, we output that graph is not Berge and stop. Else, we check if there is a near-cleaner for a shortest odd hole (\Cref{AmenableHoles}).

@@

					\mIf{$s_1 = b_1$} \Comment{calculate all possible $P_1$ paths}
						\ls $P_1(b_1) \gets$ the one-vertex path with vertex $b_1$
						\ls \algorithmicforeach ~$m \in M$ \algorithmicdo~ $P_1(m) \gets$ \textsc{undefined}
					\mElse
						\ls $P_1(b_1) \gets $ \textsc{undefined}
						\mForEach{$m \in M$}
							\mIf{ $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$ \AND
								\lsx $S_1(m)$ and $T_1(m)$ both exist \AND
								\lsx $V(S_1(m) \cap T_1(m)) = \{m\}$ \AND
								\lsx there are no edges between $V(S_1(m) \setminus m)$ 
								\lsx and $V(T_1(m) \setminus m)$ 
								\lsx
							}
								\ls $P_1(m) \gets s_1-S_1(m)-m-T_1(m)-b_1$
							\mElse
								\ls $P_1(m) \gets$ \textsc{undefined}
							\mEndIf
						\mEndFor
					\mEndIf

				\ls assign $P_2$ and $P_3$ in a similar manner
				\ls $good\_pairs_{1,2} \gets \emptyset$ \Comment{see below for definition}
				\mForEach{$m_1 \in M \cup \{b_1\}$} \Comment{calculate good $(1, 2)$-pairs}
					\mIf{$P_1(m_1) \neq \UND$}
						\ls color black the vertices of $M$ that either belong to \label{line:pyramidColor}
						\lsx $P_1(m_1)$ or have a neighbor in $P_1(m_1)$
						\lsx color all other vertices white.
						\mForEach{$m_2 \in M \cup \{b_2\}$}
							\mIf{$P_2(m_2)$ exists and contains no black vertices \lsx} \label{line:pyramidColor2}
								\ls add $(m_1, m_2)$ to $good\_pairs_{1,2}$
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls calculate $good\_pairs_{1,3}$ and $good\_pairs_{2,3}$ in similar way
				\mForEach{triple $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$} \label{line:m1m2m3}
					\mIf{$\forall_{1\leq i < j \leq 3}$: $(m_i, m_j)$ is a good $(i, j)$-pair}
						\ls \RETURN \TRUE \label{line:pyramidTrue}
					\mEndIf
				\mEndFor
			\mEndIf \label{line:pyramidEnd}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

	Checking if there exists an appropriate $a$ takes linear time (line \ref{line:a}). Calculating paths $S_i$ and $T_i$ (lines \ref{line:pyramidSStart}-\ref{line:pyramidSEnd}) takes $O(|V|^2)$ time for each $m \in M$ and $O(|V|^3)$ in total. Similarly, it takes $O(|V|^3)$ time to calculate all $P_i$ paths.

@@

	Finally, there are $O(|V|^3)$ pairs $m_1, m_2, m_3$ and checking each takes $O(1)$ time.
\end{proof}

@@

Most obvious way to find a jewel would be to enumerate all (possibly chordal) cycles of length 5 as $v_1, \ldots v_5$, check if it has all required nonedges and if it does, try to find a path $P$ as required. This gives us a time of $O(|V|^7)$. We could speed it up to $O(|V|^6)$ with more careful algorithm, but since whole Berge recognition algorithm takes time $O(|V|^9)$ and our testing showed that time it takes to test for jewels is negligible we decided against it.

@@

% \begin{wrapfigure}[15]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/t2.tex}
	\caption{An example of a $\T_2$.}%
	\vspace{-1.5cm}
\end{wrapfigure}

@@

To prove that the existence of $\T_2$ configuration implies that the graph is not Berge, we will need the following lemma:

@@

\begin{theorem}[Theorem 6.3 of \cite{MC05}]
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item if $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item if $P$ is of length 3, let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \Cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole,
		\item if $P$ is odd with the length of at least 5, it follows from \Cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2-x-p_2-\ldots-p_{n-1}-y-v_2$.
	\end{itemize}
\end{proof}

@@

\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/t3.tex}
	\caption{An example of a $\T_3$.}%
	\vspace{-1.5cm}
\end{wrapfigure}

@@

The following algorithm with running time of $O(|V|^6)$ checks whether $G$ contains a configuration of type $\T_3$:

@@

\begin{algorithmic}[1]
	\mProcedure{Contains-T3}{$G$}
	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge \AND
		\lsx $v_1v_5, v_2v_5$ are nonedges \label{line:t3start}}
		\ls $Y \gets$ the set of all $\{v_1, v_2, v_5\}$-complete vertices.
		\mForEach{$X$ -- an anticomponent of $Y$}
			\ls $F' \gets$ maximal connected subset containing $v_5$, such that $v_1, v_2$ \label{line:t3n4start}
			\lsx have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete.
			\ls $F'' \gets$ the set of all $X$-complete vertices that have a neighbor in 
			\lsx $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$
			\ls $F \gets F' \cup F''$ \label{line:t3mid}
			\mForEach{$v_4 \in V(G) \setminus\{v_1, v_2, v_5\}$, such that $v_4$ is adjacent to $v_1$
			\lsx and not to $v_2$ and $v_5$}
				\mIf{$v_4$ has a neighbor in $F$ and a nonneighbor in $X$ \label{line:t3v4if}}
					\ls $v_6 \gets$ a neighbor of $v_4$ in $F$
					\mForEach{$v_3 \in V(G) \setminus\{v_1, v_2, v_4, v_5, v_6\}$}
						\mIf{$v_3$ is adjacent to $v_2, v_4, v_5$ and not adjacent to $v_1$ \lsx}
							\ls $P \gets$ a path from $v_6$ to $v_5$ with interior in $F'$
							\ls \RETURN \TRUE \Comment{$v_1, \ldots v_6, P, X$ is a $\T_3$}
						\mEndIf
					\mEndFor
				\mEndIf
			\mEndFor \label{line:t3n4end}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

We will skip the proof that each graph containing $\T_3$ is not Berge, as it is quite technical. See section 6.7 of \cite{MC05} for the proof.

@@

\subsection{Amenable holes.}
\label{AmenableHoles}

First, let us introduce a few new definitions.

\begin{defn}[relevant triple]
	Given a graph $G$, a triple $(a, b, c)$ of vertices is called \emph{relevant} if $a \neq b$ (but possibly $c \in \{a, b\}$) and $G[\{a,b,c\}]$ is an independent set.
\end{defn}

\begin{defn}[C-major vertices]
  Given a shortest odd hole $C$ in $G$, a node $v \in V(G) \setminus V(C)$ is $C$-major if the set of its neighbors in $C$ is not contained in any 3-node path of $C$.
\end{defn}

\begin{defn}[clean odd hole]
  An odd hole $C$ of $G$ is \emph{clean} if no vertex in $G$ is $C$-major.
\end{defn}

\begin{defn}[cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{cleaner for $C$} if $X \cap V(C) = \emptyset$ and every $C$-major vertex belongs to $X$.
\end{defn}

@@

\begin{defn}[near-cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{near-cleaner for $C$} if $X$ contains all $C$-major vertices and $X \cap V(C)$ is a subset of vertex set of some 3-node path of $C$.
\end{defn}

\begin{defn}[amenable odd hole]
  An odd hole $C$ of $G$ is \emph{amenable} if it is a shortest odd hole in $G$, it is of length at least 7 and for every anticonnected set $X$ of $C$-major vertices there is a $X$-complete edge in $C$.
\end{defn}

@@

Some pictures might be helpful. \Cref{fig:amenableOddHole} shows an example of an amenable odd hole $C = v_1, \ldots, v_7$. It is not a clean hole -- vertices $v_8$ and $v_9$ are $C$-major, and $v_8, v_9, v_1, v_2, v_3$ is an axample of a near-cleaner for $C$.
\Cref{fig:notAmenableHole} shows an odd hole $C = v_1, \ldots, v_7$ that is not amenable, because there is no $\{v_8, v_9\}$-complete edge in $C$.

@@

The proof of this theorem is quite technical and we will not discuss it here. See section 8 of \cite{MC05} for the proof. We can see an example of the contraposition of the \Cref{thm:amenableHoles} working in \Cref{fig:notAmenableHole} -- vertices $v_1$, $v_2$, $v_8$, $v_6$, $v_7$ and a path $v_1-v_9-v_5-v_6$ form a jewel.

@@

\begin{alg}[List possible near cleaners, 9.2 of \cite{MC05}]
	\label{alg:listNearCleaners}
	Input: A graph $G$.

@@

\begin{alg}[Test possible near cleaner, 5.1 of \cite{MC05}]
	\label{alg:testNearCleaner}
	Input: A graph $G$ containing no simple forbidden structure, and a subset $X \subseteq V(G)$.

@@

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$} \label{line:cudaStart}
				\mForEach{3-vertex path $x_1-x_2-x_3$ of $G\setminus y_1$} \label{line:3vertex}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist} \label{line:cudaWorkStard}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2) =$\lsx$min(r(x_3, y_1), r(x_3, y_2))$}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf \label{line:cudaWorkEnd}
				\mEndFor
			\mEndFor \label{line:cudaEnd}
		\mEndFor
		\ls \RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The proof of the correctness of the other possible result is more complicated, see section 4 and Theorem 5.1 of \cite{MC05}.

@@

	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying the three conditions in the \Cref{alg:testNearCleaner} and let $y_2$ and $n$ be defined as in there. We claim that $G$ contains an odd hole.

	{\makeatletter
	\let\par\@@par
	\par\parshape0
	\everypar{}\begin{wrapfigure}{r}{0.35\textwidth}
		\input{tikzpictures/nearCleanerFound.tex}
		\caption{An odd hole is found}%
		\vspace{-0.3cm}
		\end{wrapfigure}
		Let $R(x_1, y_1) = p_1-\ldots -p_n$, and let $R(x_2, y_1) = q_1-\ldots -q_{n+1}$, where $p_1 = x_1$, $p_n = q_{n+1} = y_1$, $q_1 = x_2$ and $q_n = y_2$. From the definition of $R(x_1, y_1)$ and $R(x_2, y_1)$, none of $p_2, \ldots, p_{n-1}, q_2, \ldots, q_n$ belong to $X$. Also, from the definition of $y_1$, $y_1 \notin X$.

		Let $R(x_1, y_1) = p_1-\ldots -p_n$, and let $R(x_2, y_1) = q_1-\ldots -q_{n+1}$, where $p_1 = x_1$, $p_n = q_{n+1} = y_1$, $q_1 = x_2$ and $q_n = y_2$. From the definition of $R(x_1, y_1)$ and $R(x_2, y_1)$, none of $p_2, \ldots, p_{n-1}, q_2, \ldots, q_n$ belong to $X$. Also, from the definition of $y_1$, $y_1 \notin X$.

		Since $r(x_1, y_1) = r(x_2, y_1) - 1$ and since $x_1$, $x_2$ are nonadjacent it follows that $x_2$ does not belong to $R(x_1, y_1)$ and $x_1$ does not belong to $R(x_2, y_1)$. Since $r(x_3, y_1), r(x_3, y_2) \geq n = r(x_1, y_2)$ it follows that $x_3$ does not belong to $R(x_1, y_1)$ or to $R(x_2, y_1)$, and has no neighbors in $R(x_1, y_1) \cup R(x_2, y_1)$ other than $x_1$, $x_2$. Since $r(x_1, y_2) = n$ we get that $y_2$ does not belong to $R(x_1, y_1)$.\par}%

	We claim first that the insides of paths $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertices. For suppose that there are $2 \leq i \leq n-1$ and $2 \leq j \leq n$ that $p_i = q_j$. Then the subpaths of these two paths between $p_i$, $y_1$ are both subpaths of the shortest paths, and therefore have the same length equal to $j=i+1$. So $p_1-\ldots-p_1-q_{j+1}-\ldots-q_n$ contains a path between $x_1$, $y_2$ of length $\leq n-2$, contradicting that $r(x_1, y_2) = n$. So $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertex except $y_1$.

	If there are no edges between $R(x_1, y_1) \setminus y_1$ and $R(x_2, y_1) \setminus y_1$ then the union of these two paths and a path $x_1-x_3-x_2$ form an odd hole, so the answer is correct.

>>>>

Given a graph $G$, let us consider a problem of coloring it using as few colors as possible. If $G$ contains a clique $K$ as a subgraph, we must use at least $|V(K)|$ colors to color it. This gives us a lower bound for a chromatic number $\chi(G)$ -- it is always greater or equal to the cardinality of the largest clique $\omega(G)$. The reverse is not always true, in fact we can construct a graph with no triangles and requiring arbitrarily large numbers of colors (e.g. construction by Mycielski \cite{Mycielski1955}).

Do graphs that admit coloring using only $\omega(G)$ color are "simpler" to further analyze? Not necessarily so. Given a graph $G = (V, E)$, let us construct a graph $G'$ as the union of $G$ and a clique $K_{|V|}$. We can see that indeed $\chi(G') = \omega(G') = |V|$, but it gives us no indication of the structure of $G$ or $G'$. This leads us to the hereditary definition of \emph{perfect graphs}.

@@

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003}, algorithms to solve semidefinite programs grew out of the theory of perfect graphs. We will take a look at semidefinite programs and at perfect graph's relation to the Shannon capacity in \Cref{sec:ShannonCapacity}.

@@

\begin{theorem}[Perfect graph theorem \cite{LL72}]
	\label{thm:pgt}
	A graph is perfect if and only if its complement graph is also perfect.
\end{theorem}

@@

Then, since $\alpha(H) = \omega(\overline{H})$ and $\omega(H) = \alpha(\overline{H})$ \Cref{thm:omegaalpha} implies \Cref{thm:pgt}. We skip the proof of \Cref{thm:omegaalpha}, as it is quite technical.

@@

Berge graph recognition algorithm (later called CCLSV from the names of its authors) could be divided into two parts: first we check if either $G$ or $\overline{G}$ contain any of a number of simple forbidden structures (\Cref{SimpleStructures}). If they do, we output that graph is not Berge and stop. Else, we check if there is a near-cleaner for a shortest odd hole or for shortest odd antihole (\Cref{AmenableHoles}).

@@

					\mIf{$s_1 = b_1$ \label{line:ShortestPathsStart}} \Comment{calculate all possible $P_1$ paths}
						\ls $P_1(b_1) \gets$ the one-vertex path with vertex $b_1$
						\ls \algorithmicforeach ~$m \in M$ \algorithmicdo~ $P_1(m) \gets$ \textsc{undefined}
					\mElse
						\ls $P_1(b_1) \gets $ \textsc{undefined}
						\mForEach{$m \in M$}
							\mIf{ $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$ \AND
								\lsx $S_1(m)$ and $T_1(m)$ both exist \AND
								\lsx $V(S_1(m) \cap T_1(m)) = \{m\}$ \AND
								\lsx there are no edges between $V(S_1(m) \setminus m)$ 
								\lsx and $V(T_1(m) \setminus m)$ 
								\lsx
							}
								\ls $P_1(m) \gets s_1-S_1(m)-m-T_1(m)-b_1$
							\mElse
								\ls $P_1(m) \gets$ \textsc{undefined}
							\mEndIf
						\mEndFor
					\mEndIf

				\ls assign $P_2$ and $P_3$ in a similar manner \label{line:ShortestPathsEnd}
				\ls $good\_pairs_{1,2} \gets \emptyset$ \Comment{see below for definition}
				\mForEach{$m_1 \in M \cup \{b_1\}$} \Comment{calculate good $(1, 2)$-pairs}
					\mIf{$P_1(m_1) \neq \UND$}
						\ls color black the vertices of $M$ that either belong to \label{line:pyramidColor}
						\lsx $P_1(m_1)$ or have a neighbor in $P_1(m_1)$
						\lsx color all other vertices white.
						\mForEach{$m_2 \in M \cup \{b_2\}$}
							\mIf{$P_2(m_2)$ exists and contains no black vertices \lsx} \label{line:pyramidColor2}
								\ls add $(m_1, m_2)$ to $good\_pairs_{1,2}$
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls calculate $good\_pairs_{1,3}$ and $good\_pairs_{2,3}$ in similar way
				\mForEach{triple $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$} \label{line:m1m2m3}
					\mIf{$\forall_{1\leq i < j \leq 3}$: $(m_i, m_j)$ is a good $(i, j)$-pair}
						\ls \RETURN \TRUE \label{line:pyramidTrue}
					\mEndIf
				\mEndFor
			\mEndIf \label{line:pyramidEnd}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

@@

	Checking if there exists an appropriate $a$ takes linear time (line \ref{line:a}). Calculating paths $S_i$ and $T_i$ (lines \ref{line:pyramidSStart}-\ref{line:pyramidSEnd}) takes $O(|V|^2)$ time for each $m \in M$ and $O(|V|^3)$ in total. Similarly, it takes $O(|V|^3)$ time to calculate all $P_i$ paths (lines \ref{line:ShortestPathsStart}-\ref{line:ShortestPathsEnd}).

@@

	Finally, there are $O(|V|^3)$ pairs $m_1, m_2, m_3$ and checking each takes $O(1)$ time which gives us the overall running time of $O(|V|^9)$.
\end{proof}

@@

Notice that given a jewel $v_1, \ldots, v_5, P$, if the path $P$ is odd, then $v_1$-$P$-$v_4$-$v_5$-$v_1$ is an odd hole. If the path $P$ is even, then $v_1$-$P$-$v_4$-$v_3$-$v_2$-$v_1$ is an odd hole.

Most obvious way to find a jewel is to enumerate all (possibly chordal) cycles of length 5 as $v_1, \ldots v_5$, check if it has all required nonedges and if it does, try to find a path $P$ as required. This gives us a time of $O(|V|^7)$. We could speed it up to $O(|V|^6)$ with more careful algorithm, but since whole Berge recognition algorithm takes time $O(|V|^9)$ and our testing showed that time it takes to test for jewels is negligible we decided against it.

@@

% \begin{wrapfigure}[15]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/t2.tex}
	\caption{An example of a $\T_2$.}%
	\vspace{-.5cm}
\end{wrapfigure}

@@

To prove that the existence of a configuration of type $\T_2$ implies that the graph is not Berge, we will need the following lemma:

@@

\begin{theorem}[Theorem 6.3 of \cite{MC05}]
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item if $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item if $P$ is of length 3, let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \Cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole,
		\item if $P$ is odd with the length of at least 5, it follows from \Cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2$-$x$-$p_2$-$\ldots$-$p_{n-1}$-$y$-$v_2$.
	\end{itemize}
\end{proof}

@@

\begin{wrapfigure}{r}{0.35\textwidth}
	\vspace{-1cm}
	\input{tikzpictures/t3.tex}
	\caption{An example of a $\T_3$.}%
	\vspace{-2cm}
\end{wrapfigure}

@@

% The following algorithm with running time of $O(|V|^6)$ checks whether $G$ contains a configuration of type $\T_3$:

@@

\begin{algorithmic}[1]
	\mProcedure{Contains-T3}{$G$}
	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge \AND
		\lsx $v_1v_5, v_2v_5$ are nonedges \label{line:t3start}}
		\ls $Y \gets$ the set of all $\{v_1, v_2, v_5\}$-complete vertices.
		\mForEach{$X$ -- an anticomponent of $Y$}
			\ls $F' \gets$ maximal connected subset containing $v_5$, such that $v_1, v_2$ \label{line:t3n4start}
			\lsx have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete.
			\ls $F'' \gets$ the set of all $X$-complete vertices that have a neighbor in 
			\lsx $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$
			\ls $F \gets F' \cup F''$ \label{line:t3mid}
			\mForEach{$v_4 \in V(G) \setminus\{v_1, v_2, v_5\}$, such that $v_4$ is adjacent to $v_1$
			\lsx and not to $v_2$ and $v_5$}
				\mIf{$v_4$ has a neighbor in $F$ and a nonneighbor in $X$ \label{line:t3v4if}}
					\ls $v_6 \gets$ a neighbor of $v_4$ in $F$
					\mForEach{$v_3 \in V(G) \setminus\{v_1, v_2, v_4, v_5, v_6\}$}
						\mIf{$v_3$ is adjacent to $v_2, v_4, v_5$ and not adjacent to $v_1$ \lsx}
							\ls $P \gets$ a path from $v_6$ to $v_5$ with interior in $F'$
							\ls \RETURN \TRUE \Comment{$v_1, \ldots v_6, P, X$ is a $\T_3$} \label{line:t3found}
						\mEndIf
					\mEndFor
				\mEndIf
			\mEndFor \label{line:t3n4end}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

It is easy to see that when \Cref{alg:t3} reports a configuration of type $\T_3$ in line \ref{line:t3found}, there indeed is one. We will skip the proof that each graph containing a $\T_3$ is not Berge, as it is quite technical. See section 6.7 of \cite{MC05} for the proof.

@@

\subsection{Amenable holes.}
\label{AmenableHoles}
As we will see (\Cref{thm:amenableHoles}), if graphs $G$ and $\overline{G}$ contain no jewels, pyramids or configurations of type $\T_1, \T_2$ or $\T_3$, then every shortest odd hole in $G$ or $\overline{G}$ will have a special structure and we will call it amenable. This fact will help us determine if $G$ or $\overline{G}$ contain an odd hole. But first, let us introduce a few new definitions.

\begin{defn}[relevant triple]
	Given a graph $G$, a triple $(a, b, c)$ of vertices is called \emph{relevant} if and only if $a \neq b$ (but possibly $c \in \{a, b\}$) and $G[\{a,b,c\}]$ is an independent set.
\end{defn}

\begin{defn}[C-major vertices]
  Given a shortest odd hole $C$ in $G$, a vertex $v \in V(G) \setminus V(C)$ is $C$-major if and only if the set of its neighbors in $C$ is not contained in any 3-vertex path of $C$.
\end{defn}

\begin{defn}[clean odd hole]
  An odd hole $C$ of $G$ is \emph{clean} if and only if no vertex in $G$ is $C$-major.
\end{defn}

\begin{defn}[cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{cleaner for $C$} if and only if $X \cap V(C) = \emptyset$ and every $C$-major vertex belongs to $X$.
\end{defn}

@@

\begin{defn}[near-cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{near-cleaner for $C$} if and only if $X$ contains all $C$-major vertices and $X \cap V(C)$ is a subset of vertex set of some 3-vertex path of $C$.
\end{defn}

\begin{defn}[amenable odd hole]
  An odd hole $C$ of $G$ is \emph{amenable} if and only if it is a shortest odd hole in $G$, $|E(C)| \geq 7$, and for every anticonnected set $X$ of $C$-major vertices there is a $X$-complete edge in $C$.
\end{defn}

@@

Some pictures might be helpful. \Cref{fig:amenableOddHole} shows an example of an amenable odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$. It is not a clean hole -- vertices $v_8$ and $v_9$ are $C$-major, and $\{v_8, v_9, v_1, v_2, v_3\}$ is an axample of a near-cleaner for $C$.
\Cref{fig:notAmenableHole} shows an odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$ that is not amenable, because there is no $\{v_8, v_9\}$-complete edge in $C$.

@@

The proof of this theorem is quite technical and we will not discuss it here. See section 8 of \cite{MC05} for the proof. We can see an example of the contraposition of the \Cref{thm:amenableHoles} working in \Cref{fig:notAmenableHole} -- vertices $v_1$, $v_2$, $v_8$, $v_6$, $v_7$ and a path $v_1$-$v_9$-$v_5$-$v_6$ form a jewel.

@@

\begin{alg}[List possible near cleaners]
	\label{alg:listNearCleaners}
	Input: A graph $G$.

@@

\begin{alg}[Test possible near cleaner]
	\label{alg:testNearCleaner}
	Input: A graph $G$ containing no simple forbidden structure, and a subset $X \subseteq V(G)$.

@@

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$} \label{line:cudaStart}
				\mForEach{3-vertex path $x_1-x_2-x_3$ of $G\setminus y_1$} \label{line:3vertex}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist} \label{line:cudaWorkStard}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$ \label{line:definey2}
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2)$ \AND
						\lsx $r(x_2, y_1) \leq min(r(x_3, y_1), r(x_3, y_2))$ \label{line:manyconditions}}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf \label{line:cudaWorkEnd}
				\mEndFor
			\mEndFor \label{line:cudaEnd}
		\mEndFor
		\ls \RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

@@

	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying conditions in lines \ref{line:cudaWorkStard} and \ref{line:manyconditions}, let $y_2$ be defined as in line \ref{line:definey2} and let $k = r(x_2, y_1)$. We claim that $G$ contains an odd hole.

	{\makeatletter
	\let\par\@@par
	\par\parshape0
	\everypar{}\begin{wrapfigure}{r}{0.35\textwidth}
		\input{tikzpictures/nearCleanerFound.tex}
		\caption{An odd hole is found}%
		\vspace{-0.3cm}
		\end{wrapfigure}
		Let $R(x_1, y_1) = p_1$-$\ldots$-$p_k$, and let $R(x_2, y_1) = q_1$-$\ldots$-$q_{k+1}$, where $p_1 = x_1$, $p_k = q_{k+1} = y_1$, $q_1 = x_2$ and $q_k = y_2$. From the definition of $R(x_1, y_1)$ and $R(x_2, y_1)$, none of $p_2, \ldots, p_{k-1}, q_2, \ldots, q_k$ belong to $X$. Also, from the definition of $y_1$, $y_1 \notin X$.

		Since $r(x_1, y_1) = r(x_2, y_1) - 1$ and since $x_1$, $x_2$ are nonadjacent it follows that $x_2$ does not belong to $R(x_1, y_1)$ and $x_1$ does not belong to $R(x_2, y_1)$. Since $r(x_3, y_1), r(x_3, y_2) \geq k = r(x_1, y_2)$ it follows that $x_3$ does not belong to $R(x_1, y_1)$ or to $R(x_2, y_1)$, and has no neighbors in $R(x_1, y_1) \cup R(x_2, y_1)$ other than $x_1$, $x_2$. Since $r(x_1, y_2) = k$ we get that $y_2$ does not belong to $R(x_1, y_1)$.

		We claim first that the insides of paths $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertices. For suppose that there are $2 \leq i \leq k-1$ and $2 \leq j \leq k$ that $p_i = q_j$. Then the subpaths of these two paths between $p_i$, $y_1$ are both subpaths of the shortest paths, and therefore have the same length equal to $j=i+1$. So $p_1$-$\ldots$-$p_1$-$q_{j+1}$-$\ldots$-$q_k$ contains a path between $x_1$, $y_2$ of length $\leq k-2$, contradicting that $r(x_1, y_2) = k$. So $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertices except $y_1$.\par}%

	If there are no edges between $R(x_1, y_1) \setminus \{y_1\}$ and $R(x_2, y_1) \setminus \{y_1\}$ then the union of these two paths and a path $x_1$-$x_3$-$x_2$ forms an odd hole, so the answer is correct.

	Suppose that $p_iq_j$ is an edge for some $1\leq i \leq k-1$ and $1 \leq j \leq k$. We claim $i \geq j$. If $j = 1$ this is clear so let us assume $j > 1$. Then there is a path between $x_1$, $y_2$ within $\{p_1, \ldots , p_i, q_j, \ldots q_k\}$ which has length $\leq k-j+1$ and has no internal vertex in $X$ (since $j > 1$); and since $r(x_1, y_2) = k$, it follows that $k-j+i \geq k$, that is, $i \geq j$ as claimed.
